Embedded Scalars, Preferred Readings \& Prosody: {A}n Experimental Revisit

### 2015-02-03 text/paper/revision_01/paper.tex
K2 paper revisions my turn started

\title{Embedded Scalars, Preferred Readings \& Prosody: {A}n
  Experimental Revisit} 
\author{% Fabian
  % Schlotterbeck, Michael Franke and Petra Augurzky
} 
\date{}

@@

Given this plurality of hypothesized readings, the question arises:
how salient are these theoretically conceivable readings
compared to each other (if they are detectable by empirical means at
all)?  Addressing this question empirically is relevant because it
lies at the heart of a current debate about the nature of quantity
inferences and the demarkation between grammar (i.e. syntax/semantics)
and domain-general reasoning procedures (i.e. pragmatics). The controversy
is usually seen as one between two main competing theoretical positions
\citep[see][for a detailed comparison of the two]{ChemlaSingh2014:Remarks-on-the-}.
Simplifying things for now, these two types of theories make clearly different
predictions about how salient the local readings, in particular, should be.
The first position which we refer to as \mymark{traditionalism} (Chemla \& Singh’s ``Gricean approach'') holds that local readings should not be salient at all.
The second one which we call \mymark{grammaticalism}
(Chemla \& Singh’s ``grammatical approach'') holds that local readings
are unproblematically available and even at least as salient
as the literal ones (see Section \ref{sec:theories-predictions} for a more
thorough exposition of the theories and their predictions).

@@

More generally, two prerequisites have to be met in order to make use of
experimental data to decide between theoretical positions regarding the
computation of quantity inferences. Firstly, both traditionalism and grammaticalism
have to be amended with additional assumptions to derive testable
predictions about the relative salience of candidate readings. For example,
hypothezised effects of prosodic marking have to be taken into account before
we can test traditionalism empirically using sentences  (\ref{bsp:AE}) and (\ref{bsp:GE}).
We will address this extant position as the \mymark{prosodic markedness hypothesis}
and argue that, if prosody is hypothesized to play a central role, it is vital
to present target sentences auditorily or to control for effects of  “silent prosody”
that subjects may apply when reading a sentence. Secondly, the relative
salience of candidate readings has to be operationalized properly, i.e. we
have to devise methods to measure relative salience. For example, typicality
effects and logical dependencies between readings have to be controlled for
when evaluating speaker judgments in terms of the salience of candidate readings.
Both of these prerequisites are related to formulating what \citet{ChemlaSingh2014:Remarks-on-the-}
call ``linking functions'' which connect the core theories to the dependent
variables of an experiment. Without proper linking functions it is hard
(or impossible) to test traditionalism or grammaticalism empirically.

This paper presents results form an experimental study designed to deal with
the mentioned problems. The focus is on testing certain formulations of traditionalism
and grammaticalsim that are explicit enough to derive testable predictions about the
relative salience of candidate readings. For practical reasons, we choose what we consider
the most prominent of such formulations. In order to address the prosodic markedness
hypothesis, we presented sentence materials auditorily and explicitly manipulated contrastive
stress on embedded scalar items. In order operationalize the relative salience of candidate
readings, we employed an \mymark{incremental verification task} in which parts of a to-be-evaluated
picture were hidden and only revealed piece-wise at the participants’ request until they felt able
to give a binary truth-value judgement \citep[see][]{Conroy2008}. This way, we catered for
potential pictorial confounds, such as from typicality. To analyze the obtained data in terms
of relative salience of candidate readings, we furthermore employed a statistical model of
how the salience of readings guided participants judgments in the incremental verification task. We
validated this method by including control conditions testing ambiguous sentences with a
well-known preference on available readings. We argue that this method provides
categorical behavioral data that is both clearly indicative of the reading preferences
among participants and independent of any potential pictorial confounds, such as
the mentioned typicality effects.

The data obtained by this method is \emph{prima facie} problematic for both
traditionalism and grammaticalism (or at least what we consider the most
prominent versions thereof). However, we want to stress from the outset
that, while the experimental results allow us to derive restrictions on which
linking functions are conceivable for traditionalism and grammaticalism, conclusions
about the general validity of the two core theories will be limited
\citep[cf.][]{ChemlaSingh2014:Remarks-on-the-} because these hardly make falsifiable
predictions without the necessary linking functions. Contrary to the predictions
of common versions of traditionalism, we observed many answers indicative of
local readings of \es-sentences, irrespective of prosodic markedness. But also
prominent versions of grammaticalism are incongruent with the observed data.
Backed up by the \mymark{strongest meaning hypothesis}
\citep{DalrympleKanazawa1998:Reciprocal-Expr} as a
disambiguation criterion \citep[as suggested by][]{ChierchiaFox2008:The-Grammatical},
grammaticalism predicts local readings to be the preferred ones, but
we found no evidence for that. More strongly even, the preference order among
putative readings suggested by our data cannot even in principle be
explained by comparisons of logical strength among putative
readings. In sum, our data shed doubt on both traditionalism and grammaticalism
in their currently most common forms. We conclude the paper by some thoughts on
how either position could be amended in the light of our data.

The remainder of the paper is structured as
follows. Section~\ref{sec:get-know-your} elaborates on the three kinds
of relevant readings for our target
sentences. Section~\ref{sec:theories-predictions} works out the
predictions of traditionalism and grammaticalism (plus prominent linking functions) 
about availability and relative salience of readings for these
sentences. Section~\ref{sec:previous-studies} reviews related
experimental studies to the extent necessary to motivate our own
approach. Section~\ref{sec:exp} eventually contributes our
experimental design, analyses and results.

@@

There are two main theoretical positions which make different 
predictions about the readings of \as- and \es-sentences \citep[see][for
overview]{Horn2006:The-Border-Wars,Geurts2010:Quantity-Implic,Sauerland2012:The-Computation,ChemlaSingh2014:Remarks-on-the-}. We will refer to these as \mymark{traditionalism} and \mymark{grammaticalism}
and treat each one in turn.  As already mentioned in the introduction there is some
leeway in assessing the predictions for each approach depending on which of several
additional assumptions to adopt. Therefore, we will distinguish different varieties
of both of them. While these are certainly not the only possible varieties, they are,
in our perception, the most prominent ones. Focusing on exactly these thus has
practical relevance. Their predictions are summarized in Table~\ref{tab:predictions}
on page~\pageref{tab:predictions}. Below, in the discussion section
(section~\ref{sec:discussion-1}), further conceivable varieties of traditionalism
and grammaticalism will be discussed.

@@

Traditionalism gets its name from its conservative stance towards
Grice's original theory of conversational implicatures
\citep{Grice1975:Logic-and-Conve}. Many authors have defended
traditionalist positions in this sense. Of the more recent literature,
we would consider as traditionalist, among others, contributions by
\citet{Spector2006:Scalar-Implicat},
\citet{Sauerland2004:Scalar-Implicat},
\citet{Russell2006:Against-Grammat},
\citet{vanRooijSchulz:ExhaustiveInterpretation},
\citet{Geurts2010:Quantity-Implic} or
\citet{Franke2011:Quantity-Implic}.

According to Grice, conversational implicatures, of which quantity
implicatures are a special case, are to be thought of as
rationalizations of speaker behavior. Central in this reasoning is the
assumption that the speaker's behavior is efficient (if not optimal)
and goal-oriented. Usually, the assumed goal of conversation is the
cooperative exchange of relevant information from speaker to hearer.

@@

Consequently, there are at least two varieties of traditionalism that,
depending on which extra assumptions we make, yield slightly
different predictions: the \mymark{the strong variety of traditionalism}
maintains that the extra assumptions of the Gricean recipe hold
(e.g., as a matter of strong default) and therefore predicts that global
readings are preferred over literal ones; \mymark{the weak variety of
traditionalism} holds that the auxiliary assumptions are more fragile
and context dependent and thus predicts that literal readings may
often be preferred over global ones. In Table~\ref{tab:predictions}
(p.~\pageref{tab:predictions}), this prediction of weak traditionalism
is captured by a preference of literal over global readings because it
may \emph{a priori} seem implausible from the perspective of weak
traditionalism that all the necessary assumptions to derive global
readings are met. \fn{We mention for clarity that although weak traditionalism does
  not predict global readings to be generally preferred, it might still predict
  an \emph{epistemically weak implicature}, similar to the global
  reading, that the speaker is uncertain whether the stronger
  alternative is true. (See also the discussion in
  Section~\ref{sec:discussion-1})} For weak traditionalism, it is of course an empirical
task to determine under which circumstances the auxiliary assumptions
are met.

@@

\noindent Clearly, if we conjoin a literal reading of (\ref{bsp:AE})
with the negation of (\ref{bsp:AE-Alternatives-Extended}) we obtain
exactly the local reading. (Notice that the negation of
(\ref{bsp:AE-Alternatives-Extended}) entails the negation of
(\ref{bsp:AE-Alternative}), so it makes no difference (not) to add it
in this case.) Consequently, traditionalism does predict that the
local reading of \as-sentences is available if
(\ref{bsp:AE-Alternatives-Extended}) is an available alternative. We
therefore introduce another distinction: \mymark{restricted
  traditionalism} assumes that (\ref{bsp:AE-Alternatives-Extended}) is
not available and therefore predicts that the local reading for
\as-sentences is not available; in contrast, \mymark{unrestricted
  traditionalism} assumes that
(\ref{bsp:AE-Alternatives-Extended}) is available and so is the
local reading.\fn{The term traditionalism may be misleading here
because traditionalists usually assume that scalar implicatures should
only be computed when a sentence is uttered in a context in which a
logically stronger alternative is relevant \cite{ChemlaSingh2014:Remarks-on-the-}. However, the alternative
in (\ref{bsp:AE-Alternatives-Extended}) is not stronger than
(\ref{bsp:AE}). For completeness sake, we still decided to subsume
“unrestricted traditionalism” under traditionalism because, once
alternatives like (\ref{bsp:AE-Alternatives-Extended}) are acknowledged,
the local reading of \as-sentences can be derived in a
straightforward traditionalist manner \citep[see also][]{ChemlaSpector2010:Experimental-Ev}
who call this type of theory ``unrestricted globalism'').}

@@

The grammatical approach, as described so far, does not
make testable predictions about the relative salience of our three
candidate readings. In order to derive predictions about reading preferences,
a selection principle is needed that determines which of the
many readings made available by the grammar actually obtain or are
preferred. Proponents of grammaticalism are aware of this and have
suggested some tentative disambiguation
strategies. \citet{Magri2011:Another-Argumen}, for example, opts for
obligatory presence of $\exh$ with variability in the set of relevant
alternatives.  In the approach of \citet{Magri2011:Another-Argumen},
alternatives may be pruned entirely in the extreme case. In that case, all occurrences
of $\exh$ become vacuous resulting in a literal reading. However,
\citet{Magri2011:Another-Argumen} does not formulate an explicit theory
about which alternatives enter into the computation of sentence meaning
under what circumstances. Therefore, it is not clear what his approach
predicts about the relative salience of the candidate readings.

@@

Taking stock, traditionalism and grammaticalism make different
predictions about the relative salience (and thus preference) of our candidate
readings. Most notably, traditionalism does not predict local
readings for \es-sentences to be available at all while grammaticalism does.
Grammaticalism predicts that all candidate readings are available and,
employing the strongest meaning hypothesis as selection principle, also
comes with a clear prediction about preference: literal meanings are dispreferred.
All predictions are summarized in Table~\ref{tab:predictions}.

@@

\citet{GeurtsPouscoulous2009:Embedded-Implic} conducted a
picture-verification task (their Experiment~3) to find out whether
local readings of \as- and \es-sentences are available.\fn{We will
  focus here on a subset of the conditions tested by
  \citet{GeurtsPouscoulous2009:Embedded-Implic}. Actually, these
  authors did not use \es-sentences, but sentences where scalar
  \emph{some} was embedded under non-monotonic quantifier
  \emph{exactly two} (with appropriate pictures, of course). This case
  is a little more complex, but we will gloss over this here, treating
  their data, as if it had been obtained for \es-sentences.} Subjects
were presented with pictures similar to in Figure~\ref{fig:weak} and
\ref{fig:local} (pages \pageref{fig:weak} and \pageref{fig:local})
where the local reading gets a different truth-value than the literal
and the global reading. For \as-sentences, the local reading is false
for the critical picture in Figure~\ref{fig:weak} whereas the literal
and global readings are true; for \es-sentences the local reading is
true for the critical picture in Figure~\ref{fig:local}, whereas the
literal and global readings are false.

@@

\paragraph{Pictorial Effects.} It is often possible to present
different numbers of links between icons in pictures like in
Figures~\ref{fig:AS-distinguishing-pics}
and~\ref{fig:ES-distinguishing-pics}, without changing the
truth-values of relevant readings. Indeed, the pictorial material used
in previous studies differed in this respect, and that might explain
some of the differences in results. If that is so, then we cannot
ascribe the effects reported by
\citet{CliftonDube2010:Embedded-Implic} and
\citet{ChemlaSpector2010:Experimental-Ev} to the availability of local
readings with certainty. \citet{ChemlaSpector2010:Experimental-Ev}
also acknowledge the possibility that the \emph{typicality} of a
picture with respect to some meaning may affect graded truth-value
judgments. They show that graded judgments of \as-sentences differ
significantly for different pictures that agree on the truth value of
relevant readings but differ with respect to the amount of connections
between icons. \citeauthor{ChemlaSpector2010:Experimental-Ev} suggest
that typicality of the pictorial material can account for these
differences, but submit that this does not explain away the high
acceptance of pictures like Figure \ref{fig:weak}.

@@

\paragraph{Preference-related controls.} In order to test whether the
order of critical positions for different readings within a sequence
had an influence on responses, we included a second type of control
conditions.  These controls served as a test whether the incremental
verification task could possibly detect at least ordinal preference
relations among multiple candidate readings. Note that for our target
sentences, the logical entailment relations between readings always
require local readings to be evaluated at the end of a sequence. It
could thus in principle be possible that subjects never reach this
point, because they give truth-value judgements earlier, thereby
ending the trial. In that case it would be unclear whether these
decisions had been affected by a general unavailability of local
readings or by the fact that one of the earlier presented readings is
the preferred one. By including globally ambiguous structures with a
known preference over available readings, we thus tested whether
participants in our task occasionally choose dispreferred readings,
even if these were available only at a critical position following the
preferred readings. Finally, these conditions also controlled whether 
prosodic information can, in principle, shift answer patterns in the 
present task.

We therefore included globally ambiguous \emph{attachment ambiguities} 
as in (\ref{bsp:target-related}), which have been shown to exhibit 
interpretive preferences and can be disambiguated by prosodic information.

\begin{exe}
\ex \gll Der Brief ist mit Kreisen und Vierecken mit Sonnen
  verbunden. \label{bsp:target-related}\\
The letter is with circles and squares with suns connected.\\
The letter is connected with circles and squares with suns.
\begin{xlist}
  \ex \label{bsp:target-related-LC} The letter is connected with squares containing suns, and it is
    also connected with circles. \hfill{(\lc)}
  \ex \label{bsp:target-related-EC} The letter is connected with circles and squares, both of which
    are containing suns.  \hfill{(\ec)} 
\end{xlist}

\end{exe}
In attachment ambiguities involving post-nominal modification, 
an adjunct like a relative clause or a prepositional phrase 
(\acro{pp}) can be attached to one of two preceding hosts. 
For instance, in (\ref{bsp:target-related}), the
\acro{pp} \emph{with suns} can be attached to the preceding noun
\emph{squares}, resulting in the so-called \emph{late-closure} (\lc)
reading ((\ref{bsp:target-related-LC}). Alternatively, the \acro{pp}
can be attached to the whole conjunctive \acro{np} \emph{circles and
  squares}, corresponding to the \emph{early-closure} (\ec) reading
(\ref{bsp:target-related-EC}).  \lc-readings have been generally 
observed to be preferred over \ec-readings \citep[e.g.][]{Fodor98}
for an overview of different languages; but see \citet{CliftonCarlsonFrazier2002}
for counterevidence). Crucially, an \lc-preference for \acro{pp}-attachment 
structures has also been attested for German \citep[][]{KoniecznyHemforth2000}.

@@

In the following, we will report the results for targets and
preference-related controls separately. After providing the
descriptive results, we present statistical analyses using log-linear
models. We complement these with a Bayesian analysis that is sensitive
to the sequential nature of the incremental verification task and
probes specifically into the preference relations between candidate
readings.

@@

The previous regression-based analyses did not take the sequential
nature of the task into account. The following paragraphs therefore
spell out a generative Bayesian model which computes the
likelihood of answer patterns for different latent relative saliencies
of candidate readings and biases to answer early or late in a
sequence. We use the data to estimate the posterior likelihood of
these latent parameters to gain insight into the likely ordering
relations over the strength of latent salience of candidate readings.

@@

\noindent If we compare (\ref{bsp:observed-preferences}) with
Table~\ref{tab:predictions}, we quickly see that these findings are
not predicted by either traditionalism or grammaticalism. We address
each position separately in the following.

>>>>

\title{Embedded Scalars, Preferred Readings \& Prosody: {A}n
  Experimental Revisit} 
\author{% Michael Franke, Fabian Schlotterbeck and Petra Augurzky
} 
\date{}

@@

Given this plurality of hypothesized readings, the question arises:
how salient are these theoretically conceivable readings
compared to each other (if they are detectable by empirical means at
all)?  Addressing this question empirically is relevant because it
lies at the heart of a current debate about the nature of quantity
inferences. %  and the demarkation between grammar (i.e. syntax/semantics)
% and domain-general reasoning procedures (i.e. pragmatics).
The controversy is usually seen as one between two main competing
theoretical positions \citep[see][for a detailed comparison of the
two]{ChemlaSingh2014:Remarks-on-the-}.  Simplifying things for now,
these two types of theories make clearly different predictions about
how salient the local readings, in particular, should be.  The first
position which we refer to as \mymark{traditionalism} (similar to, but
not identical to Chemla \& Singh’s ``Gricean approach'') holds that
local readings should not be salient at all. The second one which we
call \mymark{grammaticalism} (Chemla \& Singh’s ``grammatical
approach'') holds that local readings are unproblematically available
and maybe even at least as salient as the literal ones (see Section
\ref{sec:theories-predictions} for a more thorough exposition of the
theories and their predictions).

@@

More generally, two prerequisites have to be met in order to make use
of experimental data to decide between theoretical positions regarding
the computation of quantity inferences. Firstly, both traditionalism
and grammaticalism have to be augmented with additional assumptions to
derive testable predictions about the relative salience of candidate
readings. For example, hypothezised effects of prosodic marking have
to be taken into account before we can test traditionalism empirically
using sentences (\ref{bsp:AE}) and (\ref{bsp:GE}).  We will address
this extant position as the \mymark{prosodic markedness hypothesis}
and argue that, if prosody is hypothesized to play a central role, it
is vital to present target sentences auditorily or to control for
effects of “silent prosody” that subjects may apply when reading a
sentence. Secondly, the relative salience of candidate readings has to
be operationalized properly, i.e., we have to devise methods to
measure relative salience. For example, typicality effects and logical
dependencies between readings have to be controlled for when
evaluating speaker judgments in terms of the salience of candidate
readings.  Both of these prerequisites are related to formulating what
\citet{ChemlaSingh2014:Remarks-on-the-} call ``auxiliary assumptions''
that are necessary to enable theoretical accounts to make predictions
about processing. Our emphasis here is not on processing accounts, but
on predictions of behavioral data --- what is sometimes, perhaps
misleadingly, called a computational-level perspective. But even so,
traditionalism and grammaticalism have to summon auxiliary assumptions
which connect the core theories to the dependent variables of a
behavioral experiment. Without proper auxiliary assumptions it is hard
(or impossible) to test traditionalism or grammaticalism empirically.

This paper presents results form an experimental study designed to
deal with the mentioned problems. The focus is on testing certain
formulations of traditionalism and grammaticalsim that are explicit
enough to derive testable predictions about the relative salience of
candidate readings. For practical reasons, we must choose auxiliary
assumptions that actually deliver \emph{ex ante} predictions, even
though this may not exhaust the space of variations on either core
theory that have been floated in the recent literature (see
Section~\ref{sec:theories-predictions} for discussion). In order to
address the prosodic markedness hypothesis, we presented sentence
materials auditorily and explicitly manipulated contrastive stress on
embedded scalar items. In order operationalize the relative salience
of candidate readings, we employed an \mymark{incremental verification
  task} in which parts of a to-be-evaluated picture were hidden and
only revealed piece-wise at the participants’ request until they felt
able to give a binary truth-value judgement
\citep[see][]{Conroy2008}. To analyze the obtained data in terms of
relative salience of candidate readings, we furthermore employed a
statistical model of how the salience of readings guided participants
judgments in the incremental verification task. We validated this
method by including control conditions testing ambiguous sentences
with a well-known preference on available readings. We argue that this
method provides categorical behavioral data that is both clearly
indicative of the reading preferences among participants and largely
independent of potential pictorial confounds, such as the mentioned
typicality effects.

The data obtained by this method is \emph{prima facie} problematic for
both traditionalism and grammaticalism, at least if augmented by the
auxiliary assumptions which we adopted here. However, we want to
stress from the outset that, while the experimental results allow us
to derive restrictions on which auxiliary assumptions are conceivable
for traditionalism and grammaticalism, definite conclusions about the
general validity of the two core theories will be limited
\citep[cf.][]{ChemlaSingh2014:Remarks-on-the-}. Bearing this in mind,
we can summarize our main findings crudely as follows. Contrary to the
predictions of common versions of traditionalism, we observed many
answers indicative of local readings of \es-sentences, irrespective of
prosodic markedness. But also prominent versions of grammaticalism are
incongruent with the observed data.  Backed up by the
\mymark{strongest meaning hypothesis}
\citep{DalrympleKanazawa1998:Reciprocal-Expr} as a disambiguation
criterion \citep[as suggested by][]{ChierchiaFox2008:The-Grammatical},
grammaticalism predicts local readings to be much more salient than we
found evidence for in our data. More strongly even, the preference
order among putative readings suggested by our data cannot even in
principle be explained by comparisons of logical strength among
putative readings. This finding is independently relevant in the light
of the fact that the strongest meaning hypothesis is frequently
summoned as a disambiguation criterion also in other domains, such as
the interpretation of reciprocals
\citep{DalrympleKanazawa1998:Reciprocal-Expr} or vague predicates
\citep{CobrerosEgre2012:Tolerant-Classi}. In sum, our data shed doubt
on both traditionalism and grammaticalism under prominent auxiliary
assumptions. We conclude the paper by some thoughts on how either
position could be amended in the light of our data.

The remainder of the paper is structured as
follows. Section~\ref{sec:get-know-your} elaborates on the three kinds
of relevant readings for our target
sentences. Section~\ref{sec:theories-predictions} works out the
predictions of traditionalism and grammaticalism (plus prominent auxiliary assumptions) 
about availability and relative salience of readings for these
sentences. Section~\ref{sec:previous-studies} reviews related
experimental studies to the extent necessary to motivate our own
approach. Section~\ref{sec:exp} eventually contributes our
experimental design, analyses and results.

@@

There are two main theoretical positions which make different
predictions about the readings of \as- and \es-sentences
\citep[see][for
overview]{Horn2006:The-Border-Wars,Geurts2010:Quantity-Implic,Sauerland2012:The-Computation,ChemlaSingh2014:Remarks-on-the-}. We
will refer to these as \mymark{traditionalism} and
\mymark{grammaticalism} and treat each one in turn.  As already
mentioned in the introduction, there is some leeway in assessing the
predictions for each approach depending on which auxiliary assumptions
to adopt. Therefore, we will distinguish different varieties of either
core theory. While there are certainly other varieties conceivable,
they are, in our perception, the most prominent ones, but, more
importantly, also the most practical ones in that they give us precise
and testable \emph{ex ante} predictions about expected behavioral data
(as opposed to \emph{ex post} explanations after inspecting the
data). The predictions of the considere core theories plus their
auxiliary assumptions are summarized in Table~\ref{tab:predictions} on
page~\pageref{tab:predictions}. Below, in the discussion section
(section~\ref{sec:discussion-1}), further conceivable varieties of
traditionalism and grammaticalism will be discussed \emph{after}
inspecting the relevant data.

@@

Traditionalism gets its name from its conservative stance towards
Grice's original theory of conversational implicatures
\citep{Grice1975:Logic-and-Conve}. Many authors have defended
traditionalist positions in this sense. Of the more recent literature,
we would consider as traditionalist, among others, contributions by
\citet{Spector2006:Scalar-Implicat},
\citet{Sauerland2004:Scalar-Implicat},
\citet{Russell2006:Against-Grammat},
\citet{vanRooijSchulz:ExhaustiveInterpretation},
\citet{Geurts2010:Quantity-Implic},
\citet{Franke2011:Quantity-Implic}, or \citet{GoodmanStuhlmuller2013:Knowledge-and-I}.

According to Grice, conversational implicatures, of which quantity
implicatures are a special case, are to be thought of as
rationalizations of speaker behavior. Central in this reasoning is the
assumption that the speaker's behavior is efficient (if not optimal)
and goal-oriented. Usually, the assumed goal of conversation is the
cooperative exchange of relevant information from speaker to
hearer.\footnote{Our conception of ``traditionalism'' is less
  stringent that \citeauthor{ChemlaSingh2014:Remarks-on-the-}'s
  ``Gricean approach.'' We allow our traditionalism to violate the
  constraint that all considered alternatives must be stronger than
  the one to be interpreted, so as to allow, for instance, global
  readings of \es-sentences where the alternative
  (\ref{bsp:GE-Alternative}) does not asymmetrically entail the
  to-be-interpreted (\ref{bsp:GE}) (see more below). We would even
  subsume under traditionalism abductive reasoning to the best
  explanation of what lexical meaning a speaker had intended to convey
  on an occasion of use in an embedded context, if such an account
  could be formulated consistently and in such a manner that it makes
  testable predictions.}

@@

Consequently, there are at least two varieties of traditionalism that
yield slightly different predictions, depending on which extra
assumptions we make about the relevant context of interpretation. If
the context (e.g., our experimental design) makes the extra
assumptions of the Gricean recipe needed to derive non-epistemic
global readings sufficiently salient, then we get the prediction that
global readings should be preferred over literal ones; otherwise the
prediction is that literal readings are preferred over global ones. We
call the former position \mymark{the strong variety of traditionalism}
and the latter the \mymark{the weak variety of traditionalism}.\fn{We
  mention for clarity that although weak traditionalism does not
  predict literal readings to be more salient than non-epistemic
  global readings, it might still predict an \emph{epistemically weak
    implicature}, similar to the global reading, that the speaker is
  uncertain whether the stronger alternative is true. (See also the
  discussion in Section~\ref{sec:discussion-1})} It is of course an
empirical task to determine when a context would favor either variety
of traditionalism. For our purposes, however, it suffices to note that
the core theory is plausibly compatible with either preference
relation: literal over global or the other way around (see also
Table\ref{tab:predictions}).

@@

\noindent Clearly, if we conjoin a literal reading of (\ref{bsp:AE})
with the negation of (\ref{bsp:AE-Alternatives-Extended}) we obtain
exactly the local reading. (Notice that the negation of
(\ref{bsp:AE-Alternatives-Extended}) entails the negation of
(\ref{bsp:AE-Alternative}), so it makes no difference (not) to add it
in this case.) Consequently, traditionalism does predict that the
local reading of \as-sentences is available if
(\ref{bsp:AE-Alternatives-Extended}) is an available alternative. We
therefore introduce another distinction: \mymark{restricted
  traditionalism} assumes that (\ref{bsp:AE-Alternatives-Extended}) is
not available and therefore predicts that the local reading for
\as-sentences is not available; in contrast, \mymark{unrestricted
  traditionalism} assumes that (\ref{bsp:AE-Alternatives-Extended}) is
available and so is the local reading.\fn{The term traditionalism may
  be misleading here because traditionalists usually assume that
  scalar implicatures should only be computed when a sentence is
  uttered in a context in which a logically stronger alternative is
  relevant \cite{ChemlaSingh2014:Remarks-on-the-}. However, the
  alternative in (\ref{bsp:AE-Alternatives-Extended}) is not stronger
  than (\ref{bsp:AE}). For completeness sake, we still decided to
  subsume “unrestricted traditionalism” under traditionalism because,
  once alternatives like (\ref{bsp:AE-Alternatives-Extended}) are
  acknowledged, the local reading of \as-sentences can be derived in a
  straightforward traditionalist manner \citep[see also][who call this
  type of theory ``unrestricted
  globalism'']{ChemlaSpector2010:Experimental-Ev}.}

@@

The grammatical approach, as described so far, does not yet make
falsifiable predictions about the relative salience of our three
candidate readings. In order to derive predictions about reading
preferences, a selection principle is needed that determines which of
the many readings made available by the grammar actually obtain or are
preferred. Proponents of grammaticalism are aware of this and have
suggested some tentative disambiguation
strategies. \citet{Magri2011:Another-Argumen}, for example, opts for
obligatory presence of $\exh$ with variability in the set of relevant
alternatives.  In the approach of \citet{Magri2011:Another-Argumen},
alternatives may be pruned entirely in the extreme case. In that case,
all occurrences of $\exh$ become vacuous resulting in a literal
reading. However, \citet{Magri2011:Another-Argumen} does not formulate
an explicit theory about which alternatives enter into the computation
of sentence meaning under what circumstances. Therefore, it is not
clear what his approach predicts about the relative salience of the
candidate readings.

@@

Taking stock, under the auxiliary assumptions adopted here,
traditionalism and grammaticalism make different predictions about the
relative salience (and thus preference) of our candidate
readings. Most notably, traditionalism does not predict local readings
for \es-sentences to be available at all while grammaticalism does.
Grammaticalism predicts that all candidate readings are available and,
employing the strongest meaning hypothesis as disambiguation
principle, also comes with a clear prediction about preference:
literal meanings are dispreferred.  All predictions are summarized in
Table~\ref{tab:predictions}.

@@

\citet{GeurtsPouscoulous2009:Embedded-Implic} conducted a
picture-verification task (their Experiment~3) to find out whether
local readings of \as- and \es-sentences are available.\fn{We will
  focus here on a subset of the conditions tested by
  \citet{GeurtsPouscoulous2009:Embedded-Implic}. Actually, these
  authors did not use \es-sentences, but sentences where scalar
  \emph{some} was embedded under non-monotonic quantifier
  \emph{exactly two} (with appropriate pictures, of course). This case
  is a little more complex, but we will gloss over this here, treating
  their data, as if it had been obtained for \es-sentences.} Subjects
were presented with pictures similar to those in Figure~\ref{fig:weak} and
\ref{fig:local} (pages \pageref{fig:weak} and \pageref{fig:local})
where the local reading gets a different truth-value than the literal
and the global reading. For \as-sentences, the local reading is false
for the critical picture in Figure~\ref{fig:weak} whereas the literal
and global readings are true; for \es-sentences the local reading is
true for the critical picture in Figure~\ref{fig:local}, whereas the
literal and global readings are false.

@@

\paragraph{Pictorial Effects.} It is often possible to present
different numbers of links between icons in pictures like in
Figures~\ref{fig:AS-distinguishing-pics}
and~\ref{fig:ES-distinguishing-pics}, without changing the
truth-values of relevant readings. Indeed, the pictorial material used
in previous studies differed in this respect, and that might explain
some of the differences in results. If that is so, then we cannot
ascribe the effects reported by
\citet{CliftonDube2010:Embedded-Implic} and
\citet{ChemlaSpector2010:Experimental-Ev} to the availability of local
readings with certainty. \citet{ChemlaSpector2010:Experimental-Ev}
also acknowledge the possibility that the \emph{typicality} of a
picture with respect to some meaning may affect graded truth-value
judgments: a sentence like \emph{Some of the 10 marbles are white} is
more typical, intuitively speaking, in a situation with 4 white
marbles than in a situation with 8 or 9 white marbles \citep[for more
on the controversial interpretation of ``typicality'' in this context
see][]{Tielvan-Tiel2014:Quantity-Matter,Cummins2014:Typicality-made,
  ,Franke2014:Typical-use-of-}. \citeauthor{ChemlaSpector2010:Experimental-Ev}
show that graded judgments of \as-sentences differ significantly for
different pictures that agree on the truth value of relevant readings
but differ with respect to the amount of connections between
icons. \citeauthor{ChemlaSpector2010:Experimental-Ev} suggest that
typicality of the pictorial material can account for these
differences, but submit that this does not explain away the high
acceptance of pictures like Figure \ref{fig:weak}.

@@

\paragraph{Preference-related controls.} In order to test whether the
order of critical positions for different readings within a sequence
had an influence on responses, we included a second type of control
conditions.  These controls also tested whether the incremental
verification task could possibly detect at least ordinal preference
relations among multiple candidate readings. Note that for our target
sentences, the logical entailment relations between readings always
require local readings to be evaluated at the end of a sequence. It
could be that subjects never reach this point, because they give
truth-value judgements earlier, thereby ending the trial. In that case
it would be unclear whether these decisions had been affected by a
general unavailability of local readings or by the fact that one of
the earlier presented readings is the preferred one. By including
globally ambiguous structures with a known preference over available
readings, we thus tested whether participants in our task occasionally
choose dispreferred readings, even if these were available only at a
critical position following the preferred readings. Finally, these
conditions also controlled whether prosodic information can, in
principle, shift answer patterns in the present task.

We therefore included sentences with \emph{attachment ambiguities} as
in (\ref{bsp:target-related}), which have been shown to exhibit
interpretive preferences and can be disambiguated by prosodic
information.\footnote{It is possible that attachment ambiguities are
  different from the ``pragmatic ambiguity'' in \as- and
  \es-sentences, so that information about the former is not ideally
  informative about the latter. We will address this worry in
  Section~\textcolor{red}{XYZ}, where we argue that this does not
  affect the main conclusion that we would like to draw from our data
  eventually.}

\begin{exe}
\ex \gll Der Brief ist mit Kreisen und Vierecken mit Sonnen
  verbunden. \label{bsp:target-related}\\
The letter is with circles and squares with suns connected.\\
The letter is connected with circles and squares with suns.
\begin{xlist}
  \ex \label{bsp:target-related-LC} The letter is connected with squares containing suns, and it is
    also connected with circles. \hfill{(\lc)}
  \ex \label{bsp:target-related-EC} The letter is connected with circles and squares, both of which
    are containing suns.  \hfill{(\ec)} 
\end{xlist}
\end{exe}
In attachment ambiguities involving post-nominal modification, an
adjunct like a relative clause or a prepositional phrase (\acro{pp})
can be attached to one of two preceding hosts.  For instance, in
(\ref{bsp:target-related}), the \acro{pp} \emph{with suns} can be
attached to the preceding noun \emph{squares}, resulting in the
so-called \emph{late-closure} (\lc) reading
((\ref{bsp:target-related-LC}). Alternatively, the \acro{pp} can be
attached to the whole conjunctive \acro{np} \emph{circles and
  squares}, corresponding to the \emph{early-closure} (\ec) reading
(\ref{bsp:target-related-EC}).  \lc-readings have been generally
observed to be preferred over \ec-readings (e.g., \cite{Fodor98} for
an overview of different languages, but see also
\cite{CliftonCarlsonFrazier2002} for counterevidence). Crucially, an
\lc-preference for \acro{pp}-attachment structures has also been
attested for German \citep[][]{KoniecznyHemforth2000}.

@@

In the following, we will report the results for target conditions and
preference-related controls separately. After providing the
descriptive results, we present statistical analyses using log-linear
models. We complement these with a Bayesian analysis that is sensitive
to the sequential nature of the incremental verification task and
probes specifically into the preference relations between candidate
readings.

@@

The previous regression-based analyses did not take the sequential
nature of the task into account. We noted a possible bias for exiting
later in a sequence, but ideally we would like to quantify how strong
this effect is and what this entails for the likelihood of local
readings. The following paragraphs therefore spell out a generative
Bayesian model which computes the likelihood of answer patterns for
different latent relative saliencies of candidate readings and biases
to answer early or late in a sequence. We use the data to estimate the
posterior likelihood of these latent parameters to gain insight into
the likely ordering relations over the strength of latent salience of
candidate readings.

@@

\noindent If we compare (\ref{bsp:observed-preferences}) with
Table~\ref{tab:predictions}, we quickly see that these findings are
not predicted by either traditionalism or grammaticalism under the
auxiliary assumptions that we considered so far. We address each
position separately in the following.
